---
title: "PML Project"
author: "Datasharer07"
date: "11/6/2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Practical Machine Learning Project

#Loading and Cleaning data

Loading of packages and data
```{r Load}
library(caret)
library(rpart)
library(randomForest)
library(ranger)


train<-read.csv("pml-training.csv", na.strings = c("NA", ""))
test<-read.csv("pml-testing.csv", na.strings = c("NA", ""))
```

Cleaning data includes cutting first seven unnecessary columns and excluding columns with many NAs.
```{r DataCleaning}
#remove unnecessary first 7 columns
train <- train[, -c(1:7)]

#remove columns with a lot of NA entries
train<-train[ , colSums(is.na(train)) <= 15000]
#quizz<-test[which(names(test) %in% names(train))] #Deletes all columns not in the training set
set.seed(1007)
```


# Model Building
I used data splitting to create a training and testing set.
```{r Slicing}
inTrain<-createDataPartition(y=train$classe,p=0.75,list=FALSE)
training<-train[inTrain,]
testing<-train[-inTrain,]
```

I tested two machine learning techniques: trees and random forests with the Ranger package.
```{r Learn}
#The Do parallel package lets your computer use multiple processors and thus is faster
library(doParallel)
cluster <- makeCluster(detectCores() - 1)
registerDoParallel(cluster)
#getDoParWorkers() If you want to test number of cores used
RandForestFit<-ranger(classe~.,data=training,write.forest = TRUE) #Ranger is a faster random forest
TreeFit<-train(classe~.,method="rpart",data=training)
stopCluster(cluster)
```


#Cross Validation
```{r Test}
testRandForestFit<-predict(RandForestFit,dat=testing)
testTreeFit<-predict(TreeFit,newdata=testing)
```
```{r Confusion}
AccRandForestFit<- confusionMatrix(testRandForestFit$predictions,testing$classe)
AccTreeFit<- confusionMatrix(testTreeFit,testing$classe)
```
#Expected out of sample error
The expected out of sample error for the two methods are:

- Random Forrest: `r 1-AccRandForestFit$overall[1]`
- Tree: `r 1-AccTreeFit$overall[1]`